---
title: "**Lesson 3: Reading DNA Sequences in R**"
author: "**Alejandro Navas González**"
date: today
format: 
  html:
    code-fold: False
    code-tools: true
    highlight-style: breezedark
    theme: spacelab
editor: visual
warning: false
---

```{r load_libraries}

library(tidyverse)
library(openxlsx)
library(Biostrings) 
library(DECIPHER)

```

# **Load DNA Sequence**

First, we create the DNA sequence using `DNAString()`

```{r load_seq}

seq <- DNAString("AGTGTCCCTG") 
print(seq) 

```

# **Length of the Sequence**

We can take a look at the length of the sequence using `length()`.

```{r seq_length}

len <- length(seq) 
print(len) 

```

# **GC Content of the Sequence**

To study the content in a particular base (or bases), such as the GC content, we could use `letterFrequency()`.

```{r seq_gc_content}

cC <- letterFrequency(seq, "C", as.prob = TRUE) 
cG <- letterFrequency(seq, "G", as.prob = TRUE) 
cA <- letterFrequency(seq, "A", as.prob = TRUE) 
cT <- letterFrequency(seq, "T", as.prob = TRUE) 
cGC <- letterFrequency(seq, "GC", as.prob = TRUE)

print(cGC)

```

# **Work with Fasta**

## **Read the Fasta File**

To read a sequence given in a FASTA file, we just need to write `readDNAStringSet()`.

```{r read_fa}

fastaFile <- readDNAStringSet("data/seq1.fa")
fastaFile

```


## **Extract the Names of the Sequences**

The name of the sequences contained in the FASTA file could be extracted using `names()`. 

```{r names_fa}

seq_name <- names(fastaFile)
print(seq_name)

```

## **Extract the Sequences**

```{r }

sequence <- paste(fastaFile)
print(sequence)

```

## **Compute the Length of the Sequence**

To examine the length of the sequences, there is the function `width()`. 

```{r}

length_fa <- width(fastaFile)
print(length_fa)

```

## **Count Nucleotide in the Sequence**

Finally, to see the nucleotide content for a given sequence, we can combine `table()` and `strsplit()` functions. 

```{r}

count <- table(strsplit(sequence, "")[[1]])
print(count)

```



