---
title: "**Lesson 4: DNA Replication and Reverse Complements in R**"
author: "**Alejandro Navas González**"
date: today
format: 
  html:
    code-fold: False
    code-tools: true
    highlight-style: breezedark
    theme: spacelab
editor: visual
warning: false
---

```{r load_libraries}

library(tidyverse)
library(openxlsx)
library(Biostrings) 
library(DECIPHER)

```

# **Load DNA Sequence**

First, we import the sequence using `DNAString()`. 

```{r seq_creation}

seq <- DNAString("ATGCCGTGGTAAAGCCTTAAG")
print(seq)

```

# **Manipulate the Sequence**

Create reverse complement of the sequence using `reverseComplement()` function.

```{r seq_reverse_completement}

rev <- reverseComplement(seq)
print(rev)

```

# **Save the Results**

Write output in a file using the `writeXStringSet()` function.

```{r}

writeXStringSet(DNAStringSet(rev),
                file = "results/output_reverse_complementary.fa")

```

# **Using a FASTA File**

We read a sequence from a FASTA file using `readDNAStringSet()`.

```{r read_fasta}

seq <- readDNAStringSet("https://raw.githubusercontent.com/pine-bio-support/DNA-sequences/main/sample.fna")
print(seq)

```

Finally, we create and print the reverse complement sequence.

```{r reverse_fasta}

rev <- reverseComplement(seq)
print(rev)

```





